Enhanced proximity effect in zigzag-shaped Majorana Josephson junctions.

### 2019-02-21 paper/snakemajoranas.tex
remove Analytical Estimate section
figure improvements
update the phase diagram figure caption

\begin{figure}[!htb]
\includegraphics[width=0.95\columnwidth]{figures/zigzag.pdf}
\caption{Setups. Figure of a straight and zigzag system, including trajectories.
\label{fig:setup}}
\end{figure}

@@

\begin{figure}[!htb]
\includegraphics[width=0.95\columnwidth]{figures/bandstructures}
\caption{Figure of the band stuctures corresponding to the system in Fig.~\ref{fig:setup} with different zigzag amplitudes.
The blue lines correspond to $\phi=0$, $B=0$ (trivial phase) and the orange lines to $\phi=\pi$, $B = \SI{1}{T}$ (topological phase).
The three subplots are for different amplitudes of the zigzag, with (a) a straight system $z_y=0$, (b) $z_y=\frac{W}{2}$, and (c) $z_y=W$, where $W=\SI{200}{\nm}$ is the junction width.
On the $x$-axis is the folded lattice momentum where $z_x=\SI{1300}{\nm}$.
Subplot (a) has a different scale on the $x$-axis for $k_x < 0$ than the other subplots and displays the unfolded band structure.
For the right-hand side of (a) ($k_x > 0$), (b), and (c), the folding is the same, such that the velocity $v=\frac{dE}{dk}$ can be compared visually.
We observe that once there are no more straight trajectories inside the junction (when $z_y=W$) the spectrum becomes most insensitive to the momentum $k_x$, and thus the $v_\textrm{F}$ decreases.
As the zigzag amplitude increases the band gap $E_\textrm{gap}$ increases by an order of magnitude.
The combination of these ensures a significant decrease of the Majorana decay length because $\xi_M \propto \frac{v_\textrm{F}}{E_\textrm{gap}}$
The parameters values are written at the bottom of Sec.~\ref{sec:setup}.\label{fig:band_stuctures}}
% TODO: mention z_x!
\end{figure}

@@

%%██████████████████████████████████████████████████████████████████████████
%%██ Analytical estimate
%%██████████████████████████████████████████████████████████████████████████
% \section{Analytical estimate of the band gap}

% \comment{To compute the effect of cutoff consider a single segment of the zigzag and obtain the spectrum analytically (this allows to use translation invariance).}
% We establish that the cutoff of long trajectories is the driving mechanism behind the order of magnitude gap improvement using an analytical approach.
% We derive the Andreev spectrum for a single (diagonal) segment of the zigzag, and define the gap of the system as the minimum of the spectrum occurring before the cutoff corresponding to the particular geometry.
% The system under consideration is thus a ribbon SNS junction with a tilted magnetic field.

% \comment{The definition of the cutoff...}

% \comment{We use a short junction approximation and neglect transverse SO.}
% We derive the spectrum by calculating the scattering matrix for the normal region and applying the Andreev bound state condition for the short junction limit\cite{beenakker1991universal, sticlet_robustness_2017}.
% For derivation of the scattering matrix, we neglect SOI in the transverse direction, which is valid for $W<l_\text{so}$.
% The result is specified in Eqs.~\eqref{eq:smatrix} to~\eqref{eq:momenta_short_junction}.

% \begin{align}
% \label{eq:smatrix}
%     S &= \left(
%     \begin{array}{rr}
%     r_{ll}&t_{rl}\\
%     t_{lr}&r_{rr}\\
%     \end{array}
%     \right) =
%     \left(
%     \begin{array}{rr}
%     \beta_+ & e^{-i q W} \beta_-\\
%     e^{-i q W} \beta_- & e^{-2 i q W} \beta_+\\
%     \end{array}
%     \right)
%     \\
%     \beta_\pm &= \left(
%     \begin{array}{rr}
%     e^{i \nu_{\arg}}\left(\omega^\pm_1 - \omega^\pm_2\right) & (\omega^\pm _1 + \omega^\pm _2)\\
%     -(\omega^\pm _1 + \omega^\pm _2) & e^{-i \nu _{\arg }} \left(\omega^\pm _2 - \omega^\pm _1\right)\\
%     \end{array}
%     \right)\\
%     \omega^\pm_j &= \frac{1}{4} \left(\gamma _{j} \pm \delta _{j}\right)
% \end{align}

% \begin{align}
%     \gamma_j &= \frac{q+i k_{j} \tan \left(\frac{W k_{j}}{2}\right)}{q-i k_{j} \tan \left(\frac{W k_{j}}{2}\right)} \\
%     \delta_j &= \frac{q-i k_{j} \cot \left(\frac{W k_{j}}{2}\right)}{q+i k_{j} \cot \left(\frac{W k_{j}}{2}\right)}\\
%     e^{i \nu_{\arg}} &= \frac{E_\text{z} e^{i \theta }-i \alpha  k_x}{\sqrt{E_\text{z}^2+\alpha  k_x \left(\alpha  k_x-2 E_\text{z} \sin (\theta )\right)}}
% \end{align}

% \begin{footnotesize}
% \begin{align}
%     q &= \left[ \frac{2 m_\ast}{\hbar ^2}\mu_s - k_x^2 \right]^\frac{1}{2}\\
%     k_1 &= \left[ \frac{2 m_\ast}{\hbar^2} \left(\mu_n-\sqrt{E_z^2-2 \alpha  E_z \sin (\theta ) k_x+\alpha ^2 k_x^2}\right) - k_x^2 \right]^\frac{1}{2}\\
%     k_2 &= \left[ \frac{2 m_\ast}{\hbar^2} \left(\mu_n+\sqrt{E_z^2-2 \alpha  E_z \sin (\theta ) k_x+\alpha ^2 k_x^2}\right) - k_x^2 \right]^\frac{1}{2}
% \label{eq:momenta_short_junction}
% \end{align}
% \end{footnotesize}

% In order to get the spectrum from the S-matrix [Eq.~\eqref{eq:smatrix}] we use the main result from Ref.~\onlinecite{beenakker1991universal} and start with a determinantal equation for the bound state energies in a SNS-junction:

% \begin{equation}
% \det\left[1+\alpha^{2}\left(E\right)r^{*}S_{e}\left(E,\bm{k}\right)rS_{e}^{*}\left(-E,-\bm{k}\right)\right]=0
% \end{equation}

% We rewrite this as a characteristic polynomial problem $\det\left[A-\lambda I\right]=0$, as

% \begin{equation}
% \det\left[r^{*}S_{e}\left(E,\bm{k}\right)rS_{e}^{*}\left(-E,-\bm{k}\right)-\frac{-1}{\alpha^{2}\left(E\right)}I\right]=0
% \end{equation}

% with
% \begin{equation}
% \lambda=-\frac{1}{\alpha^{2}\left(E\right)},\;A=r^{*}S_{e}\left(E,\bm{k}\right)rS_{e}^{*}\left(-E,-\bm{k}\right).
% \end{equation}

% Where $\lambda_i$ are the eigenvalues of $A$.
% Inverting $\alpha\left(E\right)\equiv\exp\left(-i\arccos\left(E/\Delta\right)\right)$ yields $\frac{E}{\Delta}=\frac{\alpha^{2}+1}{2\alpha}=\frac{1}{2}\left(\alpha+\alpha^{-1}\right)=\textrm{Re}(\alpha)$, where the last equality holds because $\alpha$ is unitary.
% Then, since $\alpha$ is only defined in the positive imaginary plane, the energies are $\frac{E_{i}}{\Delta}=\textrm{Re}\left(\sqrt{-1/\lambda_{i}}\right)$ where we take the root which has a positive imaginary part.
% Using this result and Eq.~\eqref{eq:smatrix}, we numerically find the Andreev spectrum.

% \comment{The lowest energy at $|k| < k_\textrm{cutoff}$ sets the band gap.}
% We cut off the spectrum at the momentum corresponding to the proposed cutoff in the zigzag, and compare it to the numerically obtained gap for a true zigzag system.

@@

\begin{figure}[!htb]
\includegraphics[width=0.95\columnwidth]{figures/wavefunctions}
\caption{Density of wavefunctions $\left|\psi\right|^2$ for sizes and geometries.
With (a) a straight system, (b) a saw-toothed zigzag system, (c) a system where the normal region is defined by lines parallel to a sinusoid, and (d) similar to (c) but with disordered edges.
Inside the figure, we indicate the Majorana length (or coherence length) $\xi_\textrm{M}$, the Majorana energy $E_\textrm{M}$ (the energy of the first excited state), and the topological energy gap $E_\textrm{gap}$.
We observe that $\xi_\textrm{M}$ for (a) is almost two orders of magnitude longer and $E_\textrm{gap}$ more than an order of magnitude smaller than for (b), (c), and (d).
The robustness of $E_\textrm{gap}$ and $\xi_\textrm{M}$ across (b), (c), and (d) means that the details of the geometry do not matter for the improvements to occur.
The length of the system is $3.5 L_x=\SI{4550}{\nm}$, the remaining parameters values are written at the bottom of Sec.~\ref{sec:setup}.\label{fig:wavefunctions}}
\end{figure}

@@

\begin{figure}[!htb]
\includegraphics[width=0.95\columnwidth]{figures/phasediagrams}
\caption{Phase diagrams of a straight system (a), (b) and zigzag system (c), (d), where (a), (c) are $E(\mu, B)$ and (b), (d) are $E(\phi, B)$.
Phase diagrams of a straight system (a), (b) and zigzag system (c), (d), where (a), (c) are $E(\mu, B)$ and (b), (d) are $E(\phi, B)$.
We compute the dispersion of the translationally invariant realizations of a straight and saw-toothed zigzag system and find $\min \left| {E(k)} \right|$ to obtain the energy gap.  % TODO: mention this in main text!
Additionally, in subplot (a) we plot the phase boundaries because they are too narrow to see.
We do this by solving a generalized eigenvalue problem to find all phase boundaries at once as described in Ref.~\onlinecite{nijholt2015orbital}.  % TODO: mention in main text
% TODO: mention distinguish between min E(K) and other algo in main text!
The parameters values are written at the bottom of Sec.~\ref{sec:setup}, except with $a=\SI{5}{\nm}$ and $L_\textrm{SC}=\SI{800}{\nm}$.
The chemical potential is $\mu=\SI{10}{\meV}$ in (b) and (d) and the superconducting phase is $\phi=\pi$ in (a) and (c).
\label{fig:phasediagrams}}
\end{figure}

\comment{We calculate the topological phase diagram using the gap size.}
Due to the large size of the system, we are unable to use full spectrum methods (like the Pfaffian) to compute the topological invariant of our system.
Instead, we calculate the energy gap by finding the absolute minimum of the spectrum $E_\textrm{gap}=\min{|E(k)|}$.
By both observing the gap closings and knowing the topological phase for the non-zigzag system, we can infer the topology of the system with zigzag modulation.
In Fig.~\ref{fig:phasediagrams}, we plot the energy gap as a function of magnetic field, chemical potential, $E_\textrm{gap}(B_x, \mu)$; and the superconducting phase difference $E_\textrm{gap}(B_x, \phi)$ for both a straight system [(a) and (b)] and a zigzag system [(c) and (d)].

>>>>

\begin{figure}[!htb]
\includegraphics[width=\columnwidth]{figures/zigzag.pdf}
\caption{Setups. Figure of a straight and zigzag system, including trajectories.
\label{fig:setup}}
\end{figure}

@@

\begin{figure}[!htb]
\includegraphics[width=\columnwidth]{figures/bandstructures}
\caption{Figure of the band stuctures corresponding to the system in Fig.~\ref{fig:setup} with different zigzag amplitudes.
The blue lines correspond to $\phi=0$, $B=0$ (trivial phase) and the orange lines to $\phi=\pi$, $B = \SI{1}{T}$ (topological phase).
The three subplots are for different amplitudes of the zigzag, with (a) a straight system $z_y=0$, (b) $z_y=\frac{W}{2}$, and (c) $z_y=W$, where $W=\SI{200}{\nm}$ is the junction width.
On the $x$-axis is the folded lattice momentum where $z_x=\SI{1300}{\nm}$.
Subplot (a) has a different scale on the $x$-axis for $k_x < 0$ than the other subplots and displays the unfolded band structure.
For the right-hand side of (a) ($k_x > 0$), (b), and (c), the folding is the same, such that the velocity $v=\frac{dE}{dk}$ can be compared visually.
We observe that once there are no more straight trajectories inside the junction (when $z_y=W$) the spectrum becomes most insensitive to the momentum $k_x$, and thus the $v_\textrm{F}$ decreases.
As the zigzag amplitude increases the band gap $E_\textrm{gap}$ increases by an order of magnitude.
The combination of these ensures a significant decrease of the Majorana decay length because $\xi_M \propto \frac{v_\textrm{F}}{E_\textrm{gap}}$
The parameters values are written at the bottom of Sec.~\ref{sec:setup}.\label{fig:band_stuctures}}
% TODO: mention z_x!
\end{figure}

@@

@@

\begin{figure}[!htb]
\includegraphics[width=\columnwidth]{figures/wavefunctions}
\caption{Density of wavefunctions $\left|\psi\right|^2$ for sizes and geometries.
With (a) a straight system, (b) a saw-toothed zigzag system, (c) a system where the normal region is defined by lines parallel to a sinusoid, and (d) similar to (c) but with disordered edges.
Inside the figure, we indicate the Majorana length (or coherence length) $\xi_\textrm{M}$, the Majorana energy $E_\textrm{M}$ (the energy of the first excited state), and the topological energy gap $E_\textrm{gap}$.
We observe that $\xi_\textrm{M}$ for (a) is almost two orders of magnitude longer and $E_\textrm{gap}$ more than an order of magnitude smaller than for (b), (c), and (d).
The robustness of $E_\textrm{gap}$ and $\xi_\textrm{M}$ across (b), (c), and (d) means that the details of the geometry do not matter for the improvements to occur.
The length of the system is $3.5 L_x=\SI{4550}{\nm}$, the remaining parameters values are written at the bottom of Sec.~\ref{sec:setup}.\label{fig:wavefunctions}}
\end{figure}

@@

\begin{figure}[!htb]
\includegraphics[width=\columnwidth]{figures/phasediagrams}
\caption{Part the wave function of a straight system (a) and a zigzag system (b) of length $L=\SI{3900}{\nm}$ but only the first period ($\SI{1300}{\nm}$) is plotted.
The phase diagrams (c) and (d) display $E_\textrm{gap}(B_x, \mu)$ at $\phi=\pi$ and (e) and (f) display $E_\textrm{gap}(B_x, \phi)$ at $\mu=\SI{10}{\meV}$, where $E_\textrm{gap}$ is the smallest gap in the dispersion relation.
Subplots (c) and (e) correspond to system (a), and (d) and (f) correspond to system (b).
The brown dot in the phase diagrams indicate the largest gap in the parameter space for the straight system and the parameter values for which the wavefunctions (a) and (b) are plotted.
Additionally, in subplot (c) we plot the phase boundaries because they are too narrow to see.
In (a) and (b) we note the Majorana size $\xi_\textrm{M}$ and $E_\textrm{gap}$, and see that both of these are more than an order of magnitude better for the zigzag system, even though these are plotted for the most optimal parameters for the straight system at $\mu=\SI{10}{\meV}$.
The remaining parameters values are written at the bottom of Sec.~\ref{sec:setup}, except with $a=\SI{5}{\nm}$ and $L_\textrm{SC}=\SI{800}{\nm}$.
\label{fig:phasediagrams}}
\end{figure}

\comment{We calculate the topological phase diagram using the gap size.}
In Fig.~\ref{fig:phasediagrams}, we plot the energy gap as a function of magnetic field, chemical potential, $E_\textrm{gap}(B_x, \mu)$; and the superconducting phase difference $E_\textrm{gap}(B_x, \phi)$ for both a straight system [(c) and (e)] and a zigzag system [(d) and (f)].
Additionally, we plot the first $\SI{1300}{\nm}$ (one zigzag period) of the wavefunction at the most optimal point in parameter space for the straight system.
For the straight system we calculate $E_\textrm{gap}$ by performing a binary search in $E$ for the energy at which the propagating modes start to appear.
Additionally, in Fig.~\ref{fig:phasediagrams}(c) we plot the phase boundaries that we find by solving a generalized eigenvalue problem as described in Ref.~\onlinecite{nijholt2015orbital}.
Due to the large size of the zigzag supercell, we are unable to use these full spectrum methods to compute these quantities.
Instead, we calculate $E_\textrm{gap}$ by finding the absolute minimum of the spectrum $E_\textrm{gap}=\min{|E(k)|}$.
By both observing the gap closings and knowing the topological phase for the non-zigzag system, we can infer the topology of the system with zigzag modulation.
